---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# bagofpatternsr

<!-- badges: start -->
<!-- badges: end -->

The goal of bagofpatternsr is to provide a simple implementation of the 'Bag of Patterns' time series classification algorithm as described by Lin et al (2012). It's based on the description at timeseriesclassification.com - there are other implementations of it in R, but some are built in Java which can be tricky to run.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("mattsq/bagofpatternsr")
```
## Example

Using it is easy! Models are fit like so:

```{r example}
library(bagofpatternsr)
data("FaceAll_TRAIN")
data("FaceAll_TEST")
model <- bagofpatternsr::bagofpatterns_knn(FaceAll_TRAIN, 
                                           target = "target",
                                           window_size = 20,
                                           verbose = FALSE)
```

Predictions on new datasets can be retrieved like so:

```{r cars}
new_preds  <- bagofpatternsr::predict_bagofpatterns_knn(model, 
                                                        newdata = FaceAll_TEST,
                                                        verbose = FALSE)
table(new_preds, FaceAll_TEST$target)

```

